Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2023-01-09 supplement.tex
clarifications in supp

While simulations of datasets of this size are rapid, computing LD statistics
is computationally burdensome. We simulated 500 1Mb regions under our best fit
single origin, continuous migration, and merger with stem migration models.
Using these simulated regions, we constructed 500 bootstrap replicate datasets
by sampling with replacement. Each of these 500 replicate sets were fit using
the models reported in the main text, and we compared log-likelihoods across
each fit (Figure~\ref{fig:refit-lls}). The correct model was recovered in the
all of the scenarios based on log-likelihoods alone
(Table~\ref{tab:confusion-matrix}). We conclude that our inferences are not
simply fitting statistical noise, and the observed differences in
log-likelihoods indicate a better recapitulation of the true history.

\begin{table}[ht]
    \caption{
        \label{tab:confusion-matrix}
        \textbf{Model choice from fits to simulated data.} We simulated three
        models, based on the best fit parameters from fits to data. Using these
        simulated datasets, we reinferred four models that we tested in our
        analyses in the main text, including simple (single origin) and complex
        models. Among 500 fits, we exclusively recover the simulated
        model, even when fitting more complex (i.e., more heavily
        parameterized) models to simpler simulated models (such as the single
        origin model).
    }
    \centering
    \begin{tabular}[t]{lcccc}
        \toprule
        Simulated model & Single origin & Cont. migration & Merger w/out stem mig. & Merger w/stem mig. \\
        \midrule
        Single origin & 500 & 0 & 0 & 0 \\
        Cont. migration & 0 & 500 & 0 & 0 \\
        Merger w/stem mig. & 0 & 0 & 0 & 500 \\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

While simulations of datasets of this size are rapid, computing LD statistics
is computationally burdensome. We simulated 500 1Mb regions under our best fit
single origin, continuous migration, and merger with stem migration models.
Using these simulated regions, we constructed 500 bootstrap replicate datasets
by sampling genomic regions with replacement. Each of these 500 replicate sets were fit using
the models reported in the main text, and we compared log-likelihoods across
each fit (Figure~\ref{fig:refit-lls}). The correct model was recovered in the
all of the scenarios based on log-likelihoods alone
(Table~\ref{tab:confusion-matrix}). We conclude that our inferences are not
simply fitting statistical noise, and the observed differences in
log-likelihoods indicate different abilities to capture signal present in the data.

\begin{table}[ht]
    \caption{
        \label{tab:confusion-matrix}
        \textbf{Model choice from fits to simulated data.} We simulated three
        models, based on the best fit parameters from fits to data. Using these
        simulated datasets, we reinferred four models that we tested in our
        analyses in the main text, including simple (single origin) and complex
        models. Among 500 fits, we exclusively recover the simulated
        model, even when fitting more complex (i.e., more heavily
        parameterized) models to simpler simulated models (such as the single
        origin model).
    }
    \centering
    \begin{tabular}[t]{lcccc}
        \toprule
        Simulated model$\downarrow$ & Single origin & Cont. migration & Merger w/out stem mig. & Merger w/stem mig. \\
        \midrule
        Single origin & 500 & 0 & 0 & 0 \\
        Cont. migration & 0 & 500 & 0 & 0 \\
        Merger w/stem mig. & 0 & 0 & 0 & 500 \\
        \bottomrule
    \end{tabular}
\end{table}
